Wikipedia Text Mining and Entropy Analysis

### 2019-02-02 cogsci/position-info.tex
everything in

\begin{abstract}
Although languages vary widely in their structure, all are vehicles for
the transmission of information. Consequently, aspects of speakers' word
choice can be understood through the lens models of optimal
communication. One prediction of these models is that speakers should
keep the density of information constant over the duration of an
utterance (Levy \& Jaeger, 2007). However, different languages have
different coding models that constraint the space of speaker's choices
(e.g.~canonical word order). We build on a mdethod developed by Yu,
Cong, Liang, \& Liu (2016) to analye the entropy curves of natural
language productions across a diverse set of languages and in diverse
written and spoken contexts. We show languages have characteristic
constraints that they impose on speaker's choices that predict
deviations from Uniform Information Density, and that cross-linguistic
variability in these deviations is predictable in part from syntactic
properties of those languages.

@@

These theories model the process of communication as transmission of
information over a noisy channel. The producer begins with an intended
meaning, packages this meaning into language, and then sends the meaning
to their intended receiver over a communcative channel. The receiver
must then decode from the signal they receive on their end of the
channel the producer's intended meaning. The problem is that the channel
is noisy, and sometimes the signal can get corrupted (e.g.~the producer
can misspeak, or the receiver can mishear). In order to maximize the
probability that the correct meaning is transmitted, these theories
predict that producers should choose linguistic messages that keep the
rate of information across words constant. The intuition is that if the
receiver misperceives a word, and that word contains most of the
information in the sentence, then the communication will have failed.
Because producers cannot predict which word a speaker will mishear,
their best strategy is spread the information evenly across all of the
words in a sentence, i.e.~maintain \emph{uniform information density}
(Genzel \& Charniak, 2002; Levy \& Jaeger, 2007).

The original evidence in Levy \& Jaeger (2007) finds that the insertion
of complementizers (e.g. ``that'') in relative clauses in English
corresponds to where neighboring words have high information content.
Similarly, Frank \& Jaeger (2008) argues that contradictions in English
such as ``you're'' do not occur when neighboring words are highly
informative. The evidence in favor of Uniform Information Density
largely been situation-specific and English-language driven, while the
hypothesis itself has been applied broadly over the past decade.
Applications include determining whether linguistic alignment takes
place (Jaeger \& Snider, 2013), Zipfian word length distributions
(Piantadosi, Tily, \& Gibson, 2011), communication efficiency (Mahowald,
Fedorenko, Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \&
Reitter, 2018) and the significance of ambiguity in language
(Piantadosi, Tily, \& Gibson, 2012), among other research.

However, other recent work has contradicted the Uniform Information
Density hypothesis. Similar to the original Levy \& Jaeger (2007), Zhan
\& Levy (2018) focuses on information distribution at particular points
in sentences. Zhan \& Levy (2018) finds that more information-rich
classifiers in Mandarin Chinese are produced when production of the
neighboring noun is difficult, not when the information content is high.
Jain, Singh, Ranjan, Rajkumar, \& Agarwal (2018) examine word order
across spoken sentences in Hindi, a freer word order language than
English, and find that information density has no significant effect on
word order.

@@

We extend and generalize Yu et al. (2016) in three ways: We confirm that
this same pattern is found in spoken English--both in written language,
and in conversational speech between adults and between parents and
their children. Thus, this entropy curve is a robust feature of English
productions. We then examine entropy curves cross-linguistically, and
show that characteristic curves vary across langauge the world's
languages. Finally, we show that this variation is predictable in part
from the structure of individual languages (i.e.~word order). Taken
together, our results suggest a refinement of the Uniform Information
Density hypothesis: speakers may structure their utterances to optimize
information density, but they must do so under the predictable
constraints of their language.

@@

In all oll of our studies, we used an adaptation of the by-word entropy
method developed by Yu et al. (2016). Given a text or speech corpus
divided into individual utterances, we partition the corpus by utterance
length in number of words. For each word position \(X\) of utterances of
length \(k\), we define \(w\) as a unique word occurring in position
\(X\). We further define \(p(w)\) as the number of times word \(w\)
occurs in position \(X\), divided by the number of total words that
occur in position \(X\) i.e.~the number of sentences of length \(k\).
This creates a probability distribution over the words occurring in
position \(X\), and computing the Shannon (1948) entropy \(H(X)\) of
this probability distribution gives the positional entropy of position
\(X\) in utterances of length \(k\).

@@

With this measure, we compute the unigram entropy at each position of
sentences of each length within the corpus. The result of this method
can be plotted for each utterance length as an \emph{entropy curve},
which can be visually compared across utterance length to observe the
how the unigram entropy changes across absolute positions in each of the
utterances. Genzel \& Charniak (2002) similarly examine a unigram
entropy measure on sentences, and found that entropy at the sentence
level increases linearly with sentence index within a corpus. Uniform
Information Density applies this uniformity of entropy rate in sentences
to all levels of speech, and so our method obtained from Yu et al.
(2016), which examines text at the word level, should find a
monotonically increasing at the word level.

@@

The British National Corpus consists of predominantly (90\%) written
language documents collected in the early 1980s and 1990s across a
variety of genres (scientific articles, newspapers, fiction, etc). It
also contains a small collection of spoken language. All together, it
contains \(\sim\) 100 million tokens. The Switchboard corpus is a
collection of \(\sim\) 2,400 telephone conversations between
unacquainted adults prompted with subjects of conversation. Switchboard
is the corpus used in Levy and Jaeger's (2007) original demonstration of
the Uniform Information Density Hypothesis.

@@

The entropy curves computed in both the British National Corpus and
Switchboard were remarkably consistent both across corpora and across
the range of sentence lenghts we analyzed (Figure
\ref{fig:read_and_plot_exp1}). Confirming Yu et al.'s (2016) findings,
we find that The positional entropy at the beginning of sentences is
low, then rises and plateaus for sentence-medial word positions before
dropping slightly in the second-to-last position and rising again.

@@

We analyzed three seperate corpora in CHILDES: The Providence Corpus
(Demuth, Culbertson, \& Alter, 2006), The Shiro corpus (Shiro, 2000),
and the Zhou Dinner Corpus (Li \& Zhou, 2015). The Providence corpus
consists of conversations between six 1--3-year-old American English
speaking children and their parents recorded in their homes. Providence
corpus recorded interactions between children between 1 and 3 years old
and their parents in the home. The Shiro Corpus consists of prompted
Spanish-language narratives individually collected from over a hundred
Venezualan schoolchildren, half from high SES backgrounds and half from
low SES backgrounds. The Zhou Dinner Corpus contains dinner
conversations between 5 to 6-year-old Mandarin speaking children and
their parents collected in Shanghai. Spanish is an Indo-European
language like English, possessing similar grammar, word order and
numerous cognate word. In contrast, Mandarin Chinese is typologically
completed unrelated to English.

We accessed the transcripts from each corpus using \texttt{childesr}, an
R-interface to a database formatted version of CHILDES (Sanchez et al.,
in press). We divided all utterances from each transcript into those
produced by target child, and those produced by all other speakers. We
then applied the same entropy curve method described above. For
Mandarin, we used pinyin transliterations of the utterances in the
corpus with demarcated word boundaries. The Chinese characters used for
writing Mandarin do not normally demarcate word boundaries by spacing
words apart, and for normal Chinese writing including spaces between
word boundaries can have a negative effect on reading times (Bai, Yan,
Liversedge, Zang, \& Rayner, 2008).

@@

These results present two important pieces of information. First, our
analysis of the Providence corpus broadly replicates the shape we found
in both the British National Corpus and in Switchboard. Thus, the
entropy curve of English appears not just in adult-adult conversation,
but even in speech produced by parents to their children, and speech
produced by very young children to their parents. This suggests that it
is a highly robust feature of English langauge, as it structures the
productions of even pre-schooled aged children.

Second, the Entropy curves of English, Spanish, and Mandarin were not
identical. None of these shapes resembled the monotonically increasing
function predicted naively from Uniform Information Density, but also
Mandarin was quite different from both English and Spanish. This
suggests that the entropy curve can vary from language to language. One
possibility is that this variation arises from typological features of
languages, such as characteristic word order. We noticed, for instance,
a relatively high density of determiners in the penultimate position of
English and Spanish utterances, which could account for the penultimate
dip in those languages. In our final analysis, we explored this
possibility directly, analyzing a large set of Wikipedia corpora from
diverse languages, and asking whether variability in their entropy
curves was related to variability in their syntactic structure.

\hypertarget{experiment-3}{%
\section{Experiment 3}\label{experiment-3}}

The UID hypothesis also applies to written communication: we expect
people to communicate information at a uniform rate through writing as
well. We use Wikipedia as a source for written data, which provides two
advantages. One, the quantity of data in Wikipedia is very large for
each language, at least several megabytes, and two, there are hundreds
of languages with their own Wikipedia corpora. We can therefore run our
entropy analysis on a much greater scale than using small spoken CHILDES
corpora and directly compare the results of the entropy analysis on each
language to one another. We expect to find linguistic features that
determine the shape of a language's entropy curve. We treated sentences
in the Wikipedia text corpora as equivalent to utterances in the
Switchboard and CHILDES corpora.

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

Using Giuseppe Attardi's Wikiextractor tool
\footnote{https://github.com/attardi/wikiextractor}, we extract text
corpora from Wikipedia by downloading a stored collection of Wikipedia
entries in each langauge and randomly selecting several thousand
articles from each Wikipedia language. Each language corpus was cleaned
and limited to sentences between six and \(50\) words. Similar to our
process for the Switchboard and CHILDES spoken corpora, we divided each
corpus by sentence length, and then computed our positional entropy
measure on each word position within each sentence length.

We computed two slope treatments of each curve. In the \emph{absolute}
treatment, with sentence length denoted as \(k\), we computed the slope
between positions \(1\) and \(2\), positions \(2\) and \(3\), positions
\(3\) and \(k-2\), positions \(k-2\) and \(k-1\) and positions \(k-1\)
and \(k\). For the short utterances appearing the CHILDES speech
corpora, the slopes between the first and second slopes on either end of
the distributions appeared to be more characteristic of the distribution
as a whole, with a plateau in the middle of the entropy curve for each
of the language corpora we examined in CHILDES.

However, sentences of length greater than ten in the Wikipedia corpora
were significantly more common than in the CHILDES. Therefore we also
computed relative slope treatments. In the \emph{relative 5} treatment,
we computed the slopes between every \(20\%\) of the relative word
positions in each sentence length, e.g.~the slope between \(0\%\) and
\(20\%\) of sentences of length \(10\) is the slope between the position
entropies at the sentence-initial word position and the position
representing the third word. When computing the word position index to
make those cuts, if the slope was not a whole number, then the closest
whole number position was used instead for slope calculation. Each
comparative slope within each treatment was averaged together between
different sentence lengths, for example all of the \(0\%\) to \(20\%\)
slopes over all sentence lengths were averaged together.

In both the absolute and relative \(5\) treatments, each language is
embedded in \(5\)-dimensional space. This permitted us to use cosine
similarity for direct comparison between the languages we pulled from
Wikipedia. To determine which phonological, morphological and syntactic
features affected the embedding of a language in the Wikipedia dataset,
we use the linguistic features in World Atlas of Language Structures
Dryer \& Haspelmath (2013). We checked the effects of individual
features on the embeddings of languages in the different treatments. We
computed pairwise cosine similarity between each pair of language
vectors within each treatment. For a subset of eight WALS features, we
used a generalized linear model to see whether the cosine similarity
between languages mattered in predicting if the languages shared the
same value for a WALS feature.

\hypertarget{results-and-analysis-1}{%
\subsection{Results and Analysis}\label{results-and-analysis-1}}

For eight features and \(45\) languages we pulled from Wikipedia, we
found that the cosine similarity between WALS features. The table below
shows the results for the generalized linear model we computed using
WALS features and the absolute treatment.

\begin{table}[tb]
\centering
\begin{tabular}{llrrrr}
  \hline
feature & term & estimate & std.error & statistic & p.value \\ 
  \hline
83A & cosine & 1.84 & 0.01 & 178.22 & 0.00 \\ 
  95A & cosine & 1.90 & 0.01 & 170.98 & 0.00 \\ 
  81A & cosine & 1.51 & 0.01 & 153.70 & 0.00 \\ 
  97A & cosine & 1.58 & 0.01 & 136.81 & 0.00 \\ 
  144A & cosine & 0.88 & 0.01 & 74.43 & 0.00 \\ 
  138A & cosine & 0.40 & 0.01 & 46.12 & 0.00 \\ 
  87A & cosine & 0.33 & 0.01 & 39.91 & 0.00 \\ 
  143A & cosine & 0.37 & 0.01 & 39.74 & 0.00 \\ 
  82A & cosine & 0.03 & 0.01 & 3.23 & 0.00 \\ 
   \hline
\end{tabular}
\end{table}

We also used the linear model approach to evaluate the effects of cosine
distance in the relative \(5\) treatment on the values of the same WALS
features and found similar outcomes.

From these results, we argue that cosine distance played some role in
the feature determination. The top four features in determining the WALS
feature value all characterize word order, which indicates that for this
subset of features and languages that the embeddings in the slope space
are related to the WALS features. We argue therefore that typological
features play a role in determining the positional entropy values for a
language. This indicates that the entropy curves are in some way
structured by the syntactic, morphological and phonological features of
a language.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

In this paper, we have extended and applied a model from Yu et al.
(2016) and derived from Genzel \& Charniak (2002) for distinguishing
entropy at each word position within sentences. We have argued that the
outcomes of this model are derived from the distribution of information
transmission in human communication. Language is not a uniform noisy
channel of communication, but each language individually represents a
different noisy channel of communication characterized in part by
typological features.

Our work complements the approach of studies such as Aylett \& Turk
(2004), where languages are characterized by a single number,
representing the rate of semantic information transfer per syllable.
This body of work attests to languages possessing different rates of
information transfer based on typological features as well. Within the
information transfer rate literature, the rate is an overall feature of
language that does not vary corpus to corpus. Similarly, in our work we
have obtained a characteristic entropy distribution for each language.
Both information rate and entropy distribution are characterized by
typological features. This indicates that the structure and rate of how
people communicate information varies cross-linguistically based on
typological features.

We initially wanted to use several hundred language corpora pulled from
Wikipedia and all \(144\) linguistic features from WALS. We considered
an computing unsupervized clusterings of the \(5\)-dimensional vectors
of our Wikipedia entropy curves, and comparing those clusterings to
clusterings of the WALS features directly using a cluster similarity
measure such as the Rand index Rand (1971). However, the problem then
arises of which combination of WALS features and how many features to
include in the clustering analysis, with \(144!\) different
combinations. Two additional problems presented themselves: one,
deciding how many clusters to use in an unsupervised clustering analysis
is an unsolved problem in machine learning; and two, not all of the
languages on Wikipedia have values inputted for their WALS features. As
a follow-up, we are considering missing-data imputation on the WALS
features in order to run the generalized linear model analysis on a
large dataset of languages.

On a more mechanical level, we expect that our entropy model will have
implications for how long people take to read individual words at
different positions of a sentence. Eye-tracking research has indicated
the effects of surprisal on fixation duration during eye-tracking
studies (Boston, Hale, Kliegl, Patil, \& Vasishth, 2008; Demberg \&
Keller, 2008; Smith \& Levy, 2013); higher surprisal of a word is on
average predictive of higher fixation duration. The \emph{wrap-up
effect} states that people process sentence-final words more slowly on
average then sentence-medial or sentence-initial words when reading, due
to integrating information to form a final understanding of the
sentence's meaning (Kuperman, Dambacher, Nuthmann, \& Kliegl, 2010;
Stowe, Kaan, Sabourin, \& Taylor, 2018). The wrap-up effect is drawn
from evidence in languages with a large final increase in their entropy
curve from our study, so we hypothesize that the supposed wrap-up effect
derives from the same source sentence-final increase in entropy curves
observed in English and other Germanic languages.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

{[}Not here yet.{]}

@@

\leavevmode\hypertarget{ref-aylett2004}{}%
Aylett, M., \& Turk, A. (2004). The smooth signal redundancy hypothesis:
A functional explanation for relationships between redundancy, prosodic
prominence, and duration in spontaneous speech. \emph{Language and
Speech}, \emph{47}(1), 31--56.

@@

\leavevmode\hypertarget{ref-boston2008}{}%
Boston, M. F., Hale, J., Kliegl, R., Patil, U., \& Vasishth, S. (2008).
Parsing costs as predictors of reading difficulty: An evaluation using
the potsdam sentence corpus. \emph{Journal of Eye Movement Research},
\emph{2}(1).

@@

\leavevmode\hypertarget{ref-demberg2008}{}%
Demberg, V., \& Keller, F. (2008). Data from eye-tracking corpora as
evidence for theories of syntactic processing complexity.
\emph{Cognition}, \emph{109}(2), 193--210.

@@

\leavevmode\hypertarget{ref-kuperman2010}{}%
Kuperman, V., Dambacher, M., Nuthmann, A., \& Kliegl, R. (2010). The
effect of word position on eye-movements in sentence and paragraph
reading. \emph{The Quarterly Journal of Experimental Psychology},
\emph{63}(9), 1838--1857.

@@

\leavevmode\hypertarget{ref-rand1971}{}%
Rand, W. M. (1971). Objective criteria for the evaluation of clustering
methods. \emph{Journal of the American Statistical Association},
\emph{66}(336), 846--850.

@@

\leavevmode\hypertarget{ref-smith2013}{}%
Smith, N. J., \& Levy, R. (2013). The effect of word predictability on
reading time is logarithmic. \emph{Cognition}, \emph{128}(3), 302--319.

\leavevmode\hypertarget{ref-stowe2018}{}%
Stowe, L. A., Kaan, E., Sabourin, L., \& Taylor, R. C. (2018). The
sentence wrap-up dogma. \emph{Cognition}, \emph{176}, 232--247.

>>>>

\begin{abstract}
Although languages vary widely in their structure, all are vehicles for
the transmission of information. Consequently, aspects of speakers' word
choice can be understood through the models of optimal communication.
One prediction of these models is that speakers should keep the density
of information constant over the duration of an utterance (Levy \&
Jaeger, 2007). However, different languages have different coding models
that constrain the space of a speaker's choices (e.g.~canonical word
order). We build on a method developed by Yu, Cong, Liang, \& Liu (2016)
to analyze the word-level entropy curves of natural language productions
across a diverse set of languages and in diverse written and spoken
contexts. We show languages impose characteristic constraints on
speaker's choices that predict deviations from Uniform Information
Density, and that cross-linguistic variability in these deviations is
predictable in part from syntactic properties of those languages.

@@

These theories model the process of communication as transmission of
information over a noisy channel. The producer begins with an intended
meaning, packages this meaning into language, and then sends the encoded
meaning to their intended receiver over a communicative channel. The
receiver must then decode the producer's intended meaning from the
signal they receive on their end of the channel. The problem is that the
channel is noisy, and sometimes the signal can get corrupted (e.g.~the
producer can misspeak, or the receiver can mishear). In order to
maximize the probability that the correct meaning is transmitted, these
theories predict that producers should choose linguistic messages that
keep the rate of information constant across words. The intuition is
that if the receiver misperceives a word, and that word contains most of
the information in the sentence, then the communication will have
failed. Because producers cannot predict which word a speaker will
mishear, their best strategy is spread the information evenly across all
of the words in a sentence, i.e.~maintain \emph{Uniform Information
Density} (Genzel \& Charniak, 2002; Levy \& Jaeger, 2007).

The evidence in favor of Uniform Information Density largely been
situation-specific and English-language driven, while the hypothesis
itself has been applied broadly over the past decade. The original
evidence in favor of Uniform Information Density in Levy \& Jaeger
(2007) finds that the insertion of complementizers (e.g. ``that'') in
relative clauses in English corresponds to where neighboring words have
high information content. Similarly, Frank \& Jaeger (2008) argues that
contractions in English such as ``you're'' do not occur when neighboring
words are highly informative. Applications of Uniform Information
Density include determining whether linguistic alignment takes place
(Jaeger \& Snider, 2013), Zipfian word length distributions (Piantadosi,
Tily, \& Gibson, 2011), communication efficiency (Mahowald, Fedorenko,
Piantadosi, \& Gibson, 2013), dialogue and turn-taking (Xu \& Reitter,
2018) and the significance of ambiguity in language (Piantadosi, Tily,
\& Gibson, 2012), among other research.

However, other recent work has contradicted the Uniform Information
Density hypothesis. Similar to the original Levy \& Jaeger (2007), Zhan
\& Levy (2018) focuses on information distribution at particular points
in sentences. Zhan \& Levy (2018) finds that more information-rich
classifiers in Mandarin Chinese are produced when production of the
neighboring noun is difficult, not when the information content is high.
Jain, Singh, Ranjan, Rajkumar, \& Agarwal (2018) examine word order
across spoken sentences in Hindi, a freer word order language than
English, and find that information density has no significant effect on
determining a Hindi's speaker's word order.

@@

We extend and generalize Yu et al. (2016) in three ways: We confirm that
this same pattern is found in spoken English--in written language, and
in conversational speech between adults and between parents and their
children. Thus, this entropy curve is a robust feature of English
productions. We then examine entropy curves cross-linguistically, and
show that characteristic curves vary across the world's languages.
Finally, we show that this variation is predictable in part from the
structure of individual languages (i.e.~word order). Taken together, our
results suggest a refinement of the Uniform Information Density
hypothesis: speakers may structure their utterances to optimize
information density, but they must do so under the predictable
constraints of their language.

@@

In all of our studies, we used an adaptation of the by-word entropy
method developed by Yu et al. (2016). Given a text or speech corpus
divided into individual sentences, we partition the corpus by sentence
length in number of words. For each word position \(X\) of sentences of
length \(k\), we define \(w\) as a unique word occurring in position
\(X\). We further define \(p(w)\) as the number of times word \(w\)
occurs in position \(X\), divided by the number of total words that
occur in position \(X\) i.e.~the number of sentences of length \(k\).
This creates a probability distribution over the words occurring in
position \(X\), and computing the Shannon (1948) entropy \(H(X)\) of
this probability distribution gives the positional entropy of position
\(X\) in sentences of length \(k\).

@@

With this measure, we compute the unigram entropy at each word position
of sentences of each length within the corpus. The result of this method
can be plotted for each length as an \emph{entropy curve}, which can be
visually compared across sentence lengths to observe the how the unigram
entropy changes across absolute positions in each of the sentences.
Genzel \& Charniak (2002) similarly examine a unigram entropy measure on
sentences, and found that entropy at the sentence level increases
linearly with sentence index within a corpus. Uniform Information
Density applies this uniformity of entropy rate in sentences to all
levels of speech, and so our method obtained from Yu et al. (2016),
which examines text at the word level, should find a monotonically
increasing affine function at the word level.

@@

The British National Corpus consists of predominantly (90\%) written
language documents collected in the early 1980s and 1990s across a
variety of genres (scientific articles, newspapers, fiction, etc). It
also contains a small collection of corpora from spoken language. All
together, it contains \(\sim\) 100 million tokens. The Switchboard
corpus is a collection of \(\sim\) 2,400 telephone conversations between
unacquainted adults prompted with subjects of conversation. Switchboard
is the corpus used in Levy and Jaeger's (2007) original demonstration of
the Uniform Information Density Hypothesis.

@@

The entropy curves computed in both the British National Corpus and
Switchboard were remarkably consistent both across corpora and across
the range of sentence lengths we analyzed (Figure
\ref{fig:read_and_plot_exp1}). Confirming Yu et al.'s (2016) findings,
we find that the positional entropy at the beginning of sentences is
low, then rises and plateaus for sentence-medial word positions before
dropping slightly in the second-to-last position and rising again.

@@

We analyzed three seperate corpora in CHILDES: The Providence Corpus
(Demuth, Culbertson, \& Alter, 2006), The Shiro corpus (Shiro, 2000),
and the Zhou Dinner Corpus (Li \& Zhou, 2015). The Providence corpus
consists of conversations between six 1--3-year-old American
English-speaking children and their parents recorded in their homes. The
Shiro Corpus consists of prompted Spanish-language narratives
individually collected from over a hundred Venezualan schoolchildren,
half from high SES backgrounds and half from low SES backgrounds. The
Zhou Dinner Corpus contains dinner conversations between 5--6-year-old
Mandarin-speaking children and their parents collected in Shanghai.
Spanish is an Indo-European language like English, possessing similar
grammar, word order and numerous cognate words. In contrast, Mandarin
Chinese is typologically unrelated to English.

We accessed the transcripts from each corpus using \texttt{childesr}, an
R-interface to a database formatted version of CHILDES (Sanchez et al.,
in press). We divided all utterances from each transcript into those
produced by target child, and those produced by all other speakers. We
then applied the same entropy curve method described above. For
Mandarin, we used pinyin transliterations of the utterances in the
corpus with demarcated word boundaries. The Chinese characters used for
writing Mandarin do not normally demarcate word boundaries by spacing
words apart, and for normal Chinese writing including spaces between
word boundaries is unnatural and can have a negative effect on reading
times (Bai, Yan, Liversedge, Zang, \& Rayner, 2008).

\hypertarget{results-1}{%
\subsection{Results}\label{results-1}}

@@

These results present two important pieces of information. First, the
Providence corpus entropy curve broadly replicates the curves found in
both the British National Corpus and in Switchboard. Thus, the entropy
curve of English appears not just in adult-adult conversation, but even
in speech produced by parents to their children, and speech produced by
very young children to their parents. This suggests that it is a highly
robust feature of English language, as it structures the productions of
even pre-schooled aged children.

Second, the entropy curves of English, Spanish, and Mandarin were not
identical. No shape resembled the affine function predicted naively from
Uniform Information Density, but also Mandarin was quite different from
both English and Spanish. This suggests that the entropy curve can vary
from language to language. One possibility is that this variation arises
from typological features of languages, such as characteristic word
order. We noticed, for instance, a relatively high density of
determiners in the penultimate position of English and Spanish
utterances, which could account their characteristic penultimate dip. In
our final analysis, we explored this possibility directly, analyzing a
large set of Wikipedia corpora from diverse languages, and asking
whether variability in their entropy curves was related to variability
in their syntactic structure.

\hypertarget{study-3}{%
\section{Study 3}\label{study-3}}

In order to understand how entropy curves are related to linguistic
structure, we turned to Wikipedia for a larger and more diverse of set
of languages.

\hypertarget{method}{%
\subsection{Method}\label{method}}

\hypertarget{data-and-analysis-2}{%
\subsubsection{Data and Analysis}\label{data-and-analysis-2}}

Wikipedia has two primary advantages: (1) Individual language corpora
are generally large, and (2) hundreds of languages are represented. To
understand how variability in entropy curves is related to the structure
of language, we used features from the World Atlas of Language
Structures (WALS, Dryer \& Haspelmath, 2013). WALS is The World Atlas of
Language Structures is a curated collection of typological features
across the world's languages. However, as the collection includes
contributions from many authors studying a heterogeneous set of topic
and languages, most features are missing for most languages. For our
sample, we sought to balance diversity of languages with coverage in
WALS. We selected \(45\) languages for which we \(8\) WALS features were
coded (Figure \ref{fig:clust_tree}). Unfortunately, these features all
reflected the order of words in a sentence in each langauge
(e.g.~whether objects come before verbs, whether adjectives come before
nouns). A consistent set of results across these features would be
consistent with an effect of order on entropy curves, but unfortunately
no other feature types are available for comparison.

We computed entropy curves separately for each language as before. In
order to understand how these curves varied across languages, we
developed a method for compressing the curves into a few representative
values. We noticed that curves tended to vary in predictable utterance
positions: The majority were flat in the middle of utterances, but had
either rising or falling slopes at the beginnings and ends. For each
language, for each utterance length \(k\) we computed 5 slopes: The
change between the 1st and 2nd position, the change between the 2nd and
3rd position, the slope between the 3rd position and the 3rd-to-last
position (\(k-1\)), the slope between the 3rd-to-last and the
2nd-to-last position (\(k-1\)), and the slope between the 2nd-to-last
and final positions. We then averaged these slopes across utterance
lengths within a language to estimate a language-typical signature. We
performed an additional analysis dividing each utterance into 6ths
evenly and computing the slopes between these 6ths. Results from this
analysis were qualitatively similar.

We treated these 5 slopes as a vector, allowing us to compute the
pairwise similarity between languages with cosine--a standard measure of
vector similarity (Landauer \& Dumais, 1997). Intuitively, vectors with
with a smaller angle between them are more similar. Figure
\ref{fig:clust_tree} shows a hierarchical clustering of these 5
dimensional vectors, demonstrating that entropy curve similarity appears
to capture typological relatedly, e.g.~Slavic languages are similar to
each-other.

We then used asked whether these pairwise similarities are related to
overlap in linguistic features. For each of the 8 features we measured,
we fit a logistic regression predicting whether two languages would have
the same value for it from their pairwise entropy curve similarity.
Overlap on all 8 features was reliably predicted by entropy curve
similarity (largest \(p = .001\)), with the amount of variance in
feature overlap predicted varying across predictors
\ref{tab:tab_features}. Order of object and verb (\(83A\)) was the
strongest predictor, explaining 16.22\% of the variance in pairwise
entropy curve similarity.

\begin{CodeChunk}
\begin{figure}[tb]

{\centering \includegraphics{figs/clust_tree-1}

\caption[A dendrogram estimated from hierarchically clustering the slopes of our 45 languages]{A dendrogram estimated from hierarchically clustering the slopes of our 45 languages}\label{fig:clust_tree}
\end{figure}
\end{CodeChunk}

\begin{table}[tb]
\centering
\begin{tabular}{lrrr}
  \hline
\# & t-value & p-value & r2 \\ 
  \hline
83A & 178.22 & 0.00 & 0.16 \\ 
  95A & 170.98 & 0.00 & 0.15 \\ 
  81A & 153.70 & 0.00 & 0.13 \\ 
  97A & 136.81 & 0.00 & 0.10 \\ 
  144A & 74.43 & 0.00 & 0.03 \\ 
  87A & 39.91 & 0.00 & 0.01 \\ 
  143A & 39.74 & 0.00 & 0.01 \\ 
  82A & 3.23 & 0.00 & 0.00 \\ 
   \hline
\end{tabular}
\caption{Capptoin here} 
\label{tab:tab_features}
\end{table}

\hypertarget{general-discussion}{%
\section{General Discussion}\label{general-discussion}}

Taken together, the results of our studies yield 3 main findings. First,
languages have characteristic entropy curves that can be found in their
productions whether written, spoken, or even spoken by children. These
curves diverge in reliabile ways from the monotonically increasing
functions predicting by a naive Uniform Information Density hypothesis.
Second, these characteristic curves vary across languages--some
languages like English have increasing curves, but others like Mandarin
are characterized by decreasing entropy of the course of an utterance.
Finally, these characteristic curves are related to structural
properties of those languages, for instance their word orders. This work
presents an important rejoinder to models of linguistic productions as
optimizing communicative success. While speakers may do their best to
structure the information in their productions to minimize spikes in
entropy, they operate under coding models they inherit from their native
language, and these coding models constrain the choices they make.

@@

@@

@@

@@

\leavevmode\hypertarget{ref-landauer1997}{}%
Landauer, T. K., \& Dumais, S. T. (1997). A solution to plato's problem:
The latent semantic analysis theory of acquisition, induction, and
representation of knowledge. \emph{Psychological Review}, \emph{104}(2),
211.

@@

@@

### 2019-02-02 wiki_entropy/wiki_clustering.Rmd
everything in

models %>%
  mutate(coeffs = map(model, tidy)) %>%
  select(-data, -model) %>%
  unnest() %>%
  filter(term == "cosine") %>%
  arrange(desc(statistic))
```

>>>>

models %>%
  mutate(coeffs = map(model, tidy),
         rs = map(model, )) %>%
  select(-data, -model) %>%
  unnest() %>%
  filter(term == "cosine") %>%
  arrange(desc(statistic))
```
